---
title: "R Notebook"
output:
  html_document:
    toc: yes
    toc_float: true
---

# Title level 1

## Title level 2

**bold**

*italics*

Load packages
```{r}
library(tidyverse)
library(plotly)
```



I downloaded the file and loaded it into R
```{r}
download.file("https://raw.githubusercontent.com/swcarpentry/r-novice-gapminder/gh-pages/_episodes_rmd/data/gapminder-FiveYearData.csv", destfile = "data/gapminder-FiveYearData.csv")

gapminder <- read.csv("data/gapminder-FiveYearData.csv")

head(gapminder)
```

I wonder if rstats increases life expectancy over the years


```{r}
p <- ggplot(data=gapminder,aes(x=year,y=lifeExp)) +
    geom_point()

p
```

Let's see the interactive version

```{r}
ggplotly(p)
```

# Making your own functions

If you are repeating yourself in your code, you may be able to solve that problem by making your own function!

# Data manipulation with `dplyr`

You will likely want to get subsections of your dataframe and/or calculate means of a variable for a certain subsection, `dplyr` is your friend!

# Data manipulation with `tidyr`

R likes to have 'long' format data where every row is an observation and you have a single column for 'observations' the others serve to identify that observation. (exceptions apply when you have multiple types of observations) To switch back and forth from 'wide' (how we typically enter data in a spreadsheet) to 'long' use `tidyr`


