---
title: "Software Carpentry Workshop 2017"
output:
  html_document:
    toc: yes
    toc_float: true
---

changes test test ...... . . . . not sure if this will work!

here is a new line for the internet

Load packages
```{r}

library(tidyverse)
source("functions.R")

```

I downloaded the file into R 
```{r}

download.file("https://raw.githubusercontent.com/swcarpentry/r-novice-gapminder/gh-pages/_episodes_rmd/data/gapminder-FiveYearData.csv", destfile = "data/gapminder-FiveYearData.csv")

gapminder <- read.csv("data/gapminder-FiveYearData.csv")
```
```{r}
head(gapminder)
```


I wonder what is the life expectancy over the years

```{r}
ggplot(data = gapminder, aes(x=year,y=lifeExp))+geom_point()
```

interactive version
```{r}
if(!require("plotly")){install.packages("plotly")}
library(plotly)

p <- ggplot(data = gapminder[gapminder$continent=="Americas",], aes(x=gdpPercap,y=lifeExp,color=continent,by=country))+
  geom_point()+
  scale_x_log10()+
  geom_smooth(method = "lm")+
  facet_wrap(~country)+
  xlab("Gross Domestic Product")+
  ylab("Life Expectancy")

ggplotly(p)
```

# Making your own functions

 If you are repeating yourself in your code, you may be able to solve that problem by making your own function!
 
 
```{r}

cars <- c(3,4,5,6,7,10)
se(cars)
```

  
# Data manipulation with `dplyr`
  
  You will likely want to get subsections of your dataframe and/or calculate means of a variable for a certain subsection, `dplyr` is your friend!
  
Explore select

```{r}
year_country_gdp <- select(gapminder,year,country,gdpPercap)
year_country_gdp <- select(gapminder,-pop,-continent,-lifeExp)
names(year_country_gdp)
```

Explore filter

```{r}
year_country_gdp_euro <- gapminder %>% 
  filter(continent=="Europe") %>% 
  select(year,country,gdpPercap)

euro <- filter(gapminder,continent=="Europe")
year_country_gdp_euro <- select(euro,year,country,gdpPercap)

```

Exploring the amazing group_by and summarize functions

```{r}
mean_gdp_percountry <- gapminder %>%
  group_by(country) %>% 
  summarise(mean_gdp=mean(gdpPercap),
            se_gdp=se(gdpPercap))

mean_gdp_percountry


mean_lifeExp_percontinent <- gapminder %>%
  group_by(continent,country) %>% 
  summarise(mean_lifeExp=mean(lifeExp),
            se_lifeExp=se(lifeExp),
            length_lifeExp=n())

mean_lifeExp_percontinent

```

Combining ggplot and dplyr

```{r}
euro_countries <- gapminder %>% 
  filter(continent=="Europe") %>% 
  ggplot(aes(x=year,y=lifeExp,color=country)) +
  geom_line()+
  facet_wrap(~country)

euro_countries
ggsave("euro.png")
write.csv(mean_gdp_percountry,"processed/mean_gdp_percountry.csv")

```

# Data manipulation with `tidyr`
  
R likes to have 'long' format data where every row is an observation and you have a single column for 'observations' the others serve to identify that observation. (exceptions apply when you have multiple types of observations) To switch back and forth from 'wide' (how we typically enter data in a spreadsheet) to 'long' use `tidyr`
  
```{r}

# command to download the 'wide' data
download.file("https://raw.githubusercontent.com/swcarpentry/r-novice-gapminder/gh-pages/data/gapminder_wide.csv", destfile = "data/gapminder_wide.csv")

gapminder_wide <- read.csv("data/gapminder_wide.csv")

gap_long <- gapminder_wide %>% 
  gather(obstype_year,obs_values,
         starts_with('pop'),starts_with('lifeExp'),starts_with('gdpPercap'))

gap_long <- gapminder_wide %>% 
  gather(obstype_year,obs_values,
         3:38)

```


Separate the obs_type column


```{r}
gap_normal <- gap_long %>%
  separate(obstype_year,into=c("obs_type","year"),sep="_") %>% 
  spread(obs_type,obs_values)

head(gap_normal)

gap_normal <- gap_normal %>% 
  arrange(country,continent,year)

all.equal(gapminder,gap_normal)
```


